ethan_analysis

### 2022-12-08 ethan_analysis.Rmd
Add files via upload

```{r}
#install.packages("tidymodels")
#install.packages("fs")
#install.packages("caret")
#install.packages("caTools")
#install.packages("leaps")
#install.packages("glmnet")
#install.packages("pls")
#install.packages("forecast")
```
```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}

setwd("~/Desktop/DSCI/DSCI445/Project")
library(tidyverse)
library(ggplot2)
library(dplyr)
library(caret)
library(leaps)
library(caTools)
library(tidymodels)
library(glmnet)
library(pls)
library(forecast)

@@

dataTemp <- read.csv("breast_cancer_diagnostic_data.csv")
dataTot <- dataTemp %>% select_if(~ !any(is.na(.)))
testTot <- read.csv("test.csv")
trainingTot <- read.csv("training.csv")

@@

#Dividing the data, test, and training data between malignant and benign
```{r}
#Malignant
tempM <- dataTot[dataTot$diagnosis == 'M', ]
dataM <- tempM[, !names(tempM) %in% c("diagnosis")]
splitM <- sample(c(rep(0, 0.6 * nrow(dataM)), rep(1, 0.4 * nrow(dataM))))
trainingM <- dataM[splitM == 0, ]
testM <- dataM[splitM == 1, ]

#Benign
tempB <- dataTot[dataTot$diagnosis == 'B', ]
dataB <- tempB[, !names(tempB) %in% c("diagnosis")]
splitB <- sample(c(rep(0, 0.6 * nrow(dataB)), rep(1, 0.4 * nrow(dataB))))
trainingB <- dataB[splitB == 0, ]
testB <- dataB[splitB == 1, ]
```

#Predicting the mean of the radius of the cancer cells between both Malignant & Benign combined and seperate
```{r}
#The prediction on the Total testing data set
reg_trainTot <- lm(radius_mean~., data = trainingTot)
pred_testTot <- predict(reg_trainTot, newdata = testTot)
error_testTot <- mean((testTot[, "radius_mean"] - pred_testTot)^2)

#The prediction on the Malignant testing data set
reg_trainM <- lm(radius_mean~., data = trainingM)
pred_testM <- predict(reg_trainM, newdata = testM)
error_testM <- mean((testM[, "radius_mean"] - pred_testM)^2)

#The prediction on the Benign testing data set
reg_trainB <- lm(radius_mean~., data = trainingB)
pred_testB <- predict(reg_trainB, newdata = testB)
error_testB <- mean((testB[, "radius_mean"] - pred_testB)^2)

print("Total test set error:")
print(error_testTot)
print('Malignant test set error:')
print(error_testM)
print('Benign test set error:')
print(error_testB)
```

#ridge regression
```{r}
#Total
training_matxTot <- model.matrix(radius_mean ~ . , data = trainingTot)[, -1 ]
test_matxTot <- model.matrix(radius_mean ~ . , data = testTot)[,-1]

cv_finalTot = cv.glmnet(training_matxTot , trainingTot$radius_mean, nfolds = 10, alpha=0)
lamTot = cv_finalTot$lambda.min

ridge_regTot = glmnet(training_matxTot , trainingTot$radius_mean, nfolds = 10, alpha = 0)
ridge_preTot = predict(ridge_regTot, s=lamTot, newx = test_matxTot)
error_test_ridTot <- mean((ridge_preTot - testTot$radius_mean)^2)
```
```{r}
#Malignant
training_matxM <- model.matrix(radius_mean ~ . , data = trainingM)[, -1 ]
test_matxM <- model.matrix(radius_mean ~ . , data = testM)[,-1]

cv_finalM = cv.glmnet(training_matxM , trainingM$radius_mean, nfolds = 10, alpha=0)
lamM = cv_finalM$lambda.min

ridge_regM = glmnet(training_matxM , trainingM$radius_mean, nfolds = 10, alpha = 0)
ridge_preM = predict(ridge_regM, s=lamM, newx = test_matxM)
error_test_ridM <- mean((ridge_preM - testM$radius_mean)^2)
```
```{r}
#Benign
training_matxB <- model.matrix(radius_mean ~ . , data = trainingB)[, -1 ]
test_matxB <- model.matrix(radius_mean ~ . , data = testB)[,-1]

cv_finalB = cv.glmnet(training_matxB , trainingB$radius_mean, nfolds = 10, alpha=0)
lamB = cv_finalB$lambda.min

ridge_regB = glmnet(training_matxB , trainingB$radius_mean, nfolds = 10, alpha = 0)
ridge_preB = predict(ridge_regB, s=lamB, newx = test_matxB)
error_test_ridB <- mean((ridge_preB - testB$radius_mean)^2)
```
```{r}
print("Total")
print(error_test_ridTot)
print("Malignant")
print(error_test_ridM)
print("Benign")
print(error_test_ridB)
```

#lasso
```{r}
#Total
lasso_setTot = glmnet(training_matxTot, trainingTot$radius_mean, nfolds = 10, alpha=1)
lasso_preTot = predict(lasso_setTot , s= lamTot, newx=test_matxTot)
error_test_lassoTot <- mean((lasso_preTot - testTot$radius_mean)^2)
```
```{r}
#Maligant
lasso_setM = glmnet(training_matxM, trainingM$radius_mean, nfolds = 10, alpha=1)
lasso_preM = predict(lasso_setM , s= lamM, newx=test_matxM)
error_test_lassoM <- mean((lasso_preM - testM$radius_mean)^2)
```
```{r}
#Benign
lasso_setB = glmnet(training_matxB, trainingB$radius_mean, nfolds = 10, alpha=1)
lasso_preB = predict(lasso_setB , s= lamB, newx=test_matxB)
error_test_lassoB <- mean((lasso_preB - testB$radius_mean)^2)
```
```{r}
print("Total")
print(error_test_lassoTot)
print("Malignant")
print(error_test_lassoM)
print("Benign")
print(error_test_lassoB)
```

#PCR
```{r}
#Total
trainingTot2 = dataTot %>% sample_frac(0.6)
testTot2 = dataTot %>% setdiff(trainingTot2)
pcr_fitTot2 = pcr(radius_mean ~ ., data = trainingTot2 , scale = TRUE, validation = "CV")
summary(pcr_fitTot2)
```
```{r}
#Total cont M is 31
x_trainingTot = model.matrix(radius_mean~., trainingTot2)[,-1]
x_testTot = model.matrix(radius_mean~., testTot2)[,-1]

y_trainingTot = trainingTot2 %>%
  select(radius_mean) %>%
  unlist() %>%
  as.numeric()

y_testTot = testTot2 %>%
  select(radius_mean) %>%
  unlist() %>%
  as.numeric()

pcr_predTot = predict(pcr_fitTot2, x_testTot, ncomp=31)
error_test_pcrTot <- mean((pcr_predTot-y_testTot)^2)
```

```{r}
#Malignant
trainingM2 = dataM %>% sample_frac(0.6)
testM2 = dataM %>% setdiff(trainingM2)
pcr_fitM2 = pcr(radius_mean ~ ., data = trainingM2 , scale = TRUE, validation = "CV")
summary(pcr_fitM2)
```
```{r}
#Maligant cont M is 30
x_trainingM = model.matrix(radius_mean~., trainingM2)[,-1]
x_testM = model.matrix(radius_mean~., testM2)[,-1]

y_trainingM = trainingM2 %>%
  select(radius_mean) %>%
  unlist() %>%
  as.numeric()

y_testM = testM2 %>%
  select(radius_mean) %>%
  unlist() %>%
  as.numeric()

pcr_predM = predict(pcr_fitM2, x_testM, ncomp=30)
error_test_pcrM <- mean((pcr_predM-y_testM)^2)
```

```{r}
#Belign
trainingB2 = dataB %>% sample_frac(0.6)
testB2 = dataB %>% setdiff(trainingB2)
pcr_fitB2 = pcr(radius_mean ~ ., data = trainingB2 , scale = TRUE, validation = "CV")
summary(pcr_fitB2)
```
```{r}
#Belign cont M is 29
x_trainingB = model.matrix(radius_mean~., trainingB2)[,-1]
x_testB = model.matrix(radius_mean~., testB2)[,-1]

y_trainingB = trainingB2 %>%
  select(radius_mean) %>%
  unlist() %>%
  as.numeric()

y_testB = testB2 %>%
  select(radius_mean) %>%
  unlist() %>%
  as.numeric()

pcr_predB = predict(pcr_fitB2, x_testB, ncomp=29)
error_test_pcrB <- mean((pcr_predB-y_testB)^2)
```
```{r}
print("Total")
print(error_test_pcrTot)
print("Malignant")
print(error_test_pcrM)
print("Benign")
print(error_test_pcrB)
```

#PLS
```{r}
#total
pls_fitTot = plsr(radius_mean~., data = trainingTot2, scale = TRUE, validation = "CV")
summary(pls_fitTot)
```
```{r}
#Total cont M is 30
pls_predTot = predict(pls_fitTot, x_testTot, ncomp = 30)
error_test_plsTot <- mean((pls_predTot - y_testTot)^2)
```

#Malignant
```{r}
pls_fitM = plsr(radius_mean~., data = trainingM2, scale = TRUE, validation = "CV")
summary(pls_fitM)
```
```{r}
#Malignant cont M is 30
pls_predM = predict(pls_fitM, x_testM, ncomp = 30)
error_test_plsM <- mean((pls_predM - y_testM)^2)
```

```{r}
#Benign
pls_fitB = plsr(radius_mean~., data = trainingB2, scale = TRUE, validation = "CV")
summary(pls_fitB)
```

```{r}
#Benign cont M is 26
pls_predB = predict(pls_fitB, x_testB, ncomp = 26)
error_test_plsB <- mean((pls_predB - y_testB)^2)
```
```{r}
print("Total")
print(error_test_plsTot)
print("Malignant")
print(error_test_plsM)
print("Benign")
print(error_test_plsB)
```

>>>>

```{r}
#install.packages("tidymodels")
#install.packages("fs")
#install.packages("caret")
#install.packages("caTools")
#install.packages("leaps")
#install.packages("glmnet")
#install.packages("pls")
#install.packages("forecast")
#install.packages("plyr")
#install.packages("readr")
#install.packages("repr")
```
```{r, echo=FALSE, message=FALSE, warning=FALSE, results=FALSE}

#setwd("~/Desktop/DSCI/DSCI445/Project")
library(tidyverse)
library(ggplot2)
library(dplyr)
library(caret)
library(leaps)
library(caTools)
library(tidymodels)
library(glmnet)
library(pls)
library(forecast)
library(GGally)
library(plyr)
library(readr)
library(repr)

@@

dataTemp <- read.csv("breast_cancer_diagnostic_data.csv")
datatemp2 <- dataTemp %>% select_if(~ !any(is.na(.)))
dataTot <- datatemp2[,!names(datatemp2) %in% c("id")]

testtemp <- read.csv("test.csv")
testTot <- testtemp[,!names(testtemp) %in% c("id")]

traintemp <- read.csv("training.csv")
trainTot <- traintemp[,!names(traintemp) %in% c("id")]

@@

```{r}
corr_df <- dataTot[,c('area_worst', 'area_mean', 'concave.points_mean', 'radius_worst')]
#head(corr_df)
ggpairs(corr_df)
```

--------------------------------------------------------------------------------

## area_mean
# Logestic regression

```{r}
#Test
am_lm <- lm(area_mean ~ ., data = trainTot)
pred_amtest <- predict(am_lm, testTot)
amtest_log <- postResample(pred_amtest, testTot$area_mean)
amtest_log
#summary(am_lm)
```

```{r}
#Train
pred_amtrain<- predict(am_lm, trainTot)
amtrain_log <- postResample(pred_amtrain, trainTot$area_mean)
amtrain_log
```

# ridge regression
```{r}
dummy <- dummyVars(area_mean~., data = dataTot[,-1])
x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
y_train = trainTot$area_mean

x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
y_test = testTot$area_mean

lambdas <- 10^seq(2, -3, by = -.1)
ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
optimal_lambda <- cv_ridge$lambda.min

eval_results <- function(true, predicted, df) {
  SSE <- sum((predicted - true)^2)
  SST <- sum((true - mean(true))^2)
  R_square <- 1 - SSE / SST
  RMSE = sqrt(SSE/nrow(df))

  # Model performance metrics
data.frame(
  RMSE = RMSE,
  Rsquare = R_square)

}

# Prediction and evaluation on train data
predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
amr_train <- eval_results(y_train, predictions_train, trainTot)

# Prediction and evaluation on test data
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
amr_test <- eval_results(y_test, predictions_test, testTot)

print(amr_train)
print(amr_test)
```

# lasso
```{r}
lambdas <- 10^seq(2, -3, by = -.1)
lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

lambda_best <- lasso_reg$lambda.min 
lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)

predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
aml_train <- eval_results(y_train, predictions_train, trainTot)

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
aml_test <- eval_results(y_test, predictions_test, testTot)
```

--------------------------------------------------------------------------------

## Area_worst
# Logestic regression

```{r}
#Test
aw_lm <- lm(area_worst ~ ., data = trainTot)
pred_awtest <- predict(aw_lm, testTot)
awtest_log <- postResample(pred_awtest, testTot$area_worst)
awtest_log
#summary(am_lm)
```

```{r}
#Train
pred_awtrain<- predict(aw_lm, trainTot)
awtrain_log <- postResample(pred_awtrain, trainTot$area_worst)
awtrain_log
```

# ridge regression
```{r}
dummy <- dummyVars(area_worst~., data = dataTot[,-1])
x = as.matrix(predict(dummy, newdata = trainTot[,-1]))
y_train = trainTot$area_worst

x_test = as.matrix(predict(dummy, newdata = testTot[,-1]))
y_test = testTot$area_worst

lambdas <- 10^seq(2, -3, by = -.1)
ridge_reg = glmnet(x, y_train, alpha = 0, family = 'gaussian', lambda = lambdas)
cv_ridge <- cv.glmnet(x, y_train, alpha = 0, lambda = lambdas)
optimal_lambda <- cv_ridge$lambda.min

eval_results <- function(true, predicted, df) {
  SSE <- sum((predicted - true)^2)
  SST <- sum((true - mean(true))^2)
  R_square <- 1 - SSE / SST
  RMSE = sqrt(SSE/nrow(df))

  # Model performance metrics
data.frame(
  RMSE = RMSE,
  Rsquare = R_square)

}

# Prediction and evaluation on train data
predictions_train <- predict(ridge_reg, s = optimal_lambda, newx = x)
awr_train <- eval_results(y_train, predictions_train, trainTot)

# Prediction and evaluation on test data
predictions_test <- predict(ridge_reg, s = optimal_lambda, newx = x_test)
awr_test <- eval_results(y_test, predictions_test, testTot)

print(awr_train)
print(awr_test)
```

# lasso
```{r}
lambdas <- 10^seq(2, -3, by = -.1)
lasso_reg <- cv.glmnet(x, y_train, alpha = 1, lambda = lambdas, standardize = TRUE, nfolds = 10)

lambda_best <- lasso_reg$lambda.min 
lasso_model <- glmnet(x, y_train, alpha = 1, lambda = lambda_best, standardize = TRUE)

predictions_train <- predict(lasso_model, s = lambda_best, newx = x)
awl_train <- eval_results(y_train, predictions_train, trainTot)

predictions_test <- predict(lasso_model, s = lambda_best, newx = x_test)
awl_test <- eval_results(y_test, predictions_test, testTot)
```
